### 2019-08-22 docs/rupee.tex
methods

% Template for PLoS
% Version 3.5 March 2018
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended 
% to minimize problems and delays during our production 
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file 
% and leave only the final text of your manuscript. 
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that 
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.  For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
%
% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
%
% Please add line breaks to long display equations when possible in order to fit size of the column. 
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % % 
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

%% END MACROS SECTION

% -- my additions --

\usepackage{gensymb}
\usepackage{xspace}

\newcommand{\ca}{$\alpha$-carbon\xspace}
\newcommand{\cas}{$\alpha$-carbons\xspace}

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Protein structure search to assist protein structure prediction} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Ronald Ayoub\textsuperscript{1*},
Yugyung Lee\textsuperscript{1}
\\
\bigskip
\textbf{1} School of Computing and Engineering, University of Missouri at Kansas City, Kansas City, USA
\\
\bigskip

% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
% 
% Remove or comment out the author notes below if they aren't used.
%
% Primary Equal Contribution Note
%\Yinyang These authors contributed equally to this work.

% Additional Equal Contribution Note
% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
%\ddag These authors also contributed equally to this work.

% Current address notes
%\textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
% \textcurrency b Insert second current address 
% \textcurrency c Insert third current address

% Deceased author note
%\dag Deceased

% Group/Consortium Author Note
%\textpilcrow Membership list can be found in the Acknowledgments section.

% Use the asterisk to denote corresponding authorship and provide email address in note below.
* ronaldayoub@mail.umkc.edu

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
\section*{Author summary}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

\linenumbers

% Use "Eq" instead of "Equation" for equation citations.
\section*{Introduction}

Determining the structure of a protein is an important step towards understanding its function.
Current experimental structure determination methods such as X-ray crystallography and nuclear magnetic resonance (NMR) are resource intensive and timing consuming.
Given the higher throughput of current sequencing methods, experimental protein structure determination is unable to keep pace.
Being able to predict a protein structure from its amino acid sequence would address this problem. 
However, protein structure prediction remains a central unsolved problem in molecular biology.

CASP is a biannual blind competition for protein structure prediction that began in 1994~\cite{Moult1995}.
Progress has been slow until the success of coevolutionary methods in contact prediction demonstrated in CASP11~\cite{Moult2016}.
Now, the recent success of AlphaFold at CASP13~\cite{alq2019} using deep learning combined with coevolutionary methods has refocused interest on the problem. 
While AlphaFold's performance was remarkable, the problem remains largely unsolved and more work remains to be done~\cite{alq2019}.

In~\cite{Ayoub2019} we introduced RUPEE, a purely geometric protein structure search with no dependence on sequences, clustering, or pre-calculated results. 
We compared our results with the mTM-align structure search~\cite{Dong2018}, the secondary structure matching (SSM) search~\cite{Krissinel2004}, and the CATHEDRAL structural scan~\cite{Redfern2007}, and found RUPEE is equal to or better than the best available structure searches on a benchmark of known protein domains. 
Additionally, we showed RUPEE, on average, is able to return results faster than mTM-align and CATHEDRAL.

Since the release of RUPEE, we have observed that the primary usage for RUPEE has been for uploaded protein structures that have a low maximum similarity to other structures in the PDB.
In most cases, these uploaded protein structures are the output of a protein structure prediction method. 
While in~\cite{Ayoub2019} we recognized the desirability of a purely geometric protein structure search, we did not anticipate the utility of such a search to aid in the development of protein structure prediction methods.
For instance, besides measuring the similarity of a predicted structure to a target structure in order to evaluate a protein structure prediction method in development, it is also natural to determine what known structures a predicted structure resembles most.

Recognizing the need to search on structures with low maximum similarity to known structures, we have added an additional search mode to RUPEE with increased sensitivity. 
While top-aligned mode, introduced in~\cite{Ayoub2019}, is also good for low similarity searches, the all-aligned mode we introduce below is a significant improvement over top-aligned mode in that case.

In this paper, we again compare the results of RUPEE against mTM-align, SSM, and CATHEDRAL, but this time we do so for all-aligned mode.
Furthermore, we derive a new benchmark for comparisons from protein structure predictions for free-modelling targets in CASP12 available at the CASP web site. 
Whereas in~\cite{Ayoub2019} we showed that RUPEE is equal or better than those we compared to for a benchmark of known protein structures, here we show that RUPEE, in all-aligned mode, is clearly better than those we compare to using a benchmark drawn from the structures output by protein structure prediction methods.

Originally, the RUPEE protein structure search focused on full-length matches. 
If you wanted to search for structures similar to a domain, you could search one of the protein domain classification databases such as SCOPe, CATH, or ECOD.
On the other hand, if you wanted to search for structures similar to a whole chain, you could search whole chains from the PDB. 
While the ability of RUPEE to search multiple databases for any given structure is flexible, adding the ability to execute containment searches extends this flexibility.
Therefore, RUPEE now supports explicit search types for \emph{Contained-In} and \emph{Contains} searches in addition to the \emph{Full-Length} search type.

Together, all-aligned mode and containment searches are significant additions to RUPEE that increase its utility and further demonstrates the advantage of a purely geometric protein structure search with no dependence on sequences, clustering, or pre-calculated results.

\section*{Methods}

We first give a brief outline of our linear encoding of protein structures described in more detail in~\cite{Ayoub2019}, which still remains at the core of the RUPEE protein structure search. 
Then, we describe the types of similarity scores and pairwise alignments used by RUPEE, changes to top-aligned mode, and the addition of all-aligned mode, in that order.

\subsection*{Linear encoding of protein structure}

In~\cite{Ayoub2019}, we introduced a linear encoding of protein structures.
For each pair of torsion angles combined with their secondary structure assignment, we assign an integer we refer to as a residue descriptor.
For instance, a $\beta$-turn-$\beta$ motif may be represented as a sequence of residue descriptors, similar to that shown below. 
\begin{gather}\label{E:descrseq} 
    [\, 5, 5, 5, 5, 5, 5, 7, 5, 11, 11, 5, 5, 5, 5, 5, 5 \,]
\end{gather}

Next, we derive a multiset of overlapping 3-grams of residue descriptors. 
This representation is often referred to as shingling and was developed in~\cite{Broder1997a}. 
The overlap between shingles ensures some of the order information within the original sequence is preserved in the multiset.

By shingling, we obtain a multiset of ordered lists from an ordered list of residue descriptors. 
As an example, the sequence in (\ref{E:descrseq}) becomes the following multiset of shingles. 
\begin{align}\label{E:shinglebag}
    \begin{split}
        \{\,&[5, 5, 5], [5, 5, 5], [5, 5, 5], [5, 5, 5], [5, 5, 7] \\
            & [5, 7, 5], [7, 5, 11], [5, 11, 11], [11, 11, 5], [11, 5, 5] \\
            & [5, 5, 5], [5, 5, 5], [5, 5, 5], [5, 5, 5] \,\}
    \end{split}
\end{align}

Next, each shingle $s$ is hashed to an integer as shown in~(\ref{E:hashdef}). 
\begin{gather}\label{E:hashdef}
    s_{hash} = s_1 \times 13^2 + s_2 \times 13 + s_3
\end{gather}

After hashing, the multiset in (\ref{E:shinglebag}) becomes the following multiset of integers.
\begin{align}\label{E:hashbag}
    \begin{split}
    \{\,&915, 915, 915, 915, 917, 941, 1259 \\
        &999, 2007, 1929, 915, 915, 915, 915 \,\}
    \end{split}
\end{align}

The hashing completes the transformation of an ordered list of residue descriptors to a multiset of integers that still retains some of the order information present in the original list.

Notice in (\ref{E:hashbag}) the value 915, corresponding to the shingle $[ 5, 5, 5 ]$, occurs frequently indicating the presence of $\beta$-strands. 
To address this lack of specificity, we introduced a heuristic we call \emph{run position encoding} (RPE), where a run is a consecutive sequence of identical descriptors. 
To distinguish between short and long runs, thereby increasing the specificity of the shingles, we add a factor of $10^5$ to each shingle hash as a function of the first residue's position in a run. 
Applying to the multiset of integers in (\ref{E:hashbag}) gives
\begin{align}\label{E:rpebag}
    \begin{split}
    \{\,&00915, 10915, 20915, 20915, 10917, 00941, 01259 \\
        &00999, 02007, 01929, 00915, 10915, 20915, 20915 \,\}
    \end{split}
\end{align}
where the leading zero run factors are shown for clarity.

The pyramidal approach for the run factors preserves matches at the boundaries between secondary structure runs and loops that would not otherwise be preserved in the presence of differences in run lengths of one or more.

\subsection*{Similarity scores and alignments}

For both top-aligned mode and all-aligned mode, the final step in a structure search is to run pairwise structure alignments between the query structure and each matched result and return them in sorted order. 
However, before this final step, a few different filtering methods are used to pick the candidates for pairwise structure alignments.
Here, we describe the filtering methods in order of increasing complexity and end with a description of what alignment tool RUPEE uses for pairwise structure alignments.

When representing each protein structure using a multiset of integers as shown in (\ref{E:rpebag}), full-length similarity for a candidate pair of structures $a$ and $b$ is defined using the Jaccard similarity~\cite{Levan1971} for multisets,
\begin{align}
    J(a,b) = \frac{\sum_i min(a_i, b_i)}{\sum_i max(a_i,b_i)}\text{,}
\end{align}
where $i$ ranges over all possible shingle hashes $s_i$ and $a_i$ and $b_i$ give the counts of shingle hash $s_i$ in structures $a$ and $b$, respectively. 
Determining the Jaccard similarity between two structures represented as multisets is very fast but is too coarse-grained a comparison to be of any greater use than as a first-pass filtering method.

When regarding the multiset in (\ref{E:rpebag}) as an ordered list of integers, we assign a pair of protein structures a full-length similarity score using a simplified Needleman-Wunsch~\cite{Needleman1970} sequence alignment algorithm where mismatches and gaps are penalized -1 points and matches are awarded +1 points.
For containment searches, depending on whether or not the search type is Contained-In or Contains, one of the sides of the dynamic programming matrix is not penalized for the opening gap and end gap.
This is often referred to as semi-global sequence alignment and was developed in~\cite{Brudno2003}. 
The relative order of scores for protein structure comparisons using this method is more accurate than the rough similarity estimates provided by the Jaccard similarity and are still very fast to perform. 
However, they are still only suitable for coarse-grained filtering.

Pairwise structure alignment, the most accurate method for comparing protein structures, involves finding a set of spatial rotations and translations for two protein structures that minimizes a distance metric. 
Traditionally, the root mean squared deviation (RMSD) between \cas of aligned residues is minimized.
However, the RMSD score does not factor in the distance between unaligned residues nor does it consider the percentage of aligned residues, that is, alignment \emph{coverage}. 
RMSD scores also have some dependence on the length of the aligned proteins. 
On the other hand, the TM-score~\cite{Zhang2004} takes all residues into account and normalizes for both coverage and length of the aligned proteins. 
For this reason, TM-score is frequently used in scoring protein structure alignments and accordingly, we use it for sorting the final results provided by RUPEE.

RUPEE uses TM-align~\cite{Zhang2005} for performing pairwise structure alignments.
TM-align uses a rotation matrix designed to maximize the TM-score rather than minimizing the RMSD along with dynamic programming to find the best alignment. 
In addition to using global or semi-global sequence alignment with the Needleman-Wunsch algorithm, for full-length and containment search types respectively, we apply the same logic to how normalization is used in the TM-align algorithm in order for it to be compatible with the initial filtering methods.
For the Full-Length searches we normalize by the average length of both structures, for Contained-In searches we normalize by the query structure and for Contains searches we normalize by the matched structure.
The 3 types of normalization available in TM-align and how they affect searches are discussed in greater detail in the results section of~\cite{Ayoub2019}.

RUPEE makes use of the TM-align algorithm in a few different context during a search depending on the search mode and type. 
These differents usages of TM-align will be discussed below as they occur. 
ed in more detail below in the context in which it is used.

[got here]

\subsection*{Top-aligned mode}

For top-aligned mode, RUPEE uses min-hashing~\cite{Broder1998} and locality sensitive hashing (LSH)~\cite{Indyk1998} as described in detail in our previous work~\cite{Ayoub2019} to quickly estimate the Jaccard similarity of a query protein against all available structures in the searched database.  
Then, the top scoring 40000 protein structures are obtained based on the Jaccard similarity estimates to complete the initial filtering.

After the initial filtering, we obtain more accurate similarity scores for the filtered matches using the simplified Needleman-Wunsch algorithm. 
Here, the dynamic programming matrix for Needleman-Wunsch is scored as described above based on the chosen search type, Contained-In, Contains, or Full-Length. 
Then, we execute TM-align on the top scoring 8000 matches from Needleman-Wunsch using a reduced number of dynamic programming iterations in the TM-align algorithm. 
We sort these initial alignments by TM-score and obtain the top 400 matches.
Finally, we execute TM-align using the default number of dynamic programming iterations on the top 400 matches and return the results sorted by TM-score.

In addition to using global or semi-global alignment with the Needleman-Wunsch algorithm, for full-length and containment search types respectively, we also apply the same logic to how normalization is used in the TM-align algorithm. 
For the Full-Length searches we normalize by the average length of both structures, for Contained-In searches we normalize by the query structure and for Contains searches we normalize by the matched structure.
The 3 types of normalization available in TM-align and how they affect searches are discussed in greater detail in the results section of~\cite{Ayoub2019}.

The filter sizes of 40000, 8000 and 400 have been chosen based on quality of results and speed. 
We have found that increasing the size of either of these filters results in only marginal improvements in the quality of results. 
Given that performing pairwise alignments is the most time-consuming aspect of the RUPEE structure search, the marginal improvements gained from larger filter sizes have to be balanced against the number of pairwise alignments performed.

\subsection*{All-aligned mode}

While top-aligned search mode may be sufficient for searching for known protein structures, as was shown in~\cite{Ayoub2019} and will also be reinforced below, the need for greater sensitivity arises when searching with structures output from protein structure prediction protocols, where the resultant structure may only have a maximum TM-score of less than 0.50 when compared against all available structures.
Furthermore, for top-aligned mode, the initial filtering is based on full-length matches using the Jaccard similarities, which is not suitable for containment searches. We address both of these concerns by the addition of all-aligned mode.

The first difference between all-aligned and top-aligned modes is that all-aligned modes skips the min-hashing and LSH step. 
This has the effect of reducing the scalability of all-aligned mode. 
However, we have found that with the current sizes of protein structure databases still being less than a million structures, fast alignments estimates can still be performed on the entire collection using effective multithreading and return within a reasonable amount of time, more often less than 10 minutes on structures with 300 or few residues.

[I'm not sure if distinguished between TM-align and NW algorithms above in terms of NW being much faster than TM-align.
perhaps this can be done in related work].

% Place figure captions after the first paragraph in which they are cited.
\begin{figure*}[!h]
\includegraphics{Fig5}
\caption{{\bf Bold the figure title.}
Figure caption text here, please use this space for the figure panel descriptions instead of using subfigure commands. A: Lorem ipsum dolor sit amet. B: Consectetur adipiscing elit.}
\label{fig5}
\end{figure*}

% For figure citations, please use "Fig" instead of "Figure".
Nulla mi mi, Fig~\ref{fig1} venenatis sed ipsum varius, volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, \nameref{S1_Video} vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.

% Place figure captions after the first paragraph in which they are cited.
\begin{figure*}[!h]
\includegraphics{Fig1}
\caption{{\bf Bold the figure title.}
Figure caption text here, please use this space for the figure panel descriptions instead of using subfigure commands. A: Lorem ipsum dolor sit amet. B: Consectetur adipiscing elit.}
\label{fig1}
\end{figure*}

% Results and Discussion can be combined.
\section*{Results}
Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.

% Place tables after the first paragraph in which they are cited.
\begin{table}[!ht]
\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\centering
\caption{
{\bf Table caption Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.}}
\begin{tabular}{|l+l|l|l|l|l|l|l|}
\hline
\multicolumn{4}{|l|}{\bf Heading1} & \multicolumn{4}{|l|}{\bf Heading2}\\ \thickhline
$cell1 row1$ & cell2 row 1 & cell3 row 1 & cell4 row 1 & cell5 row 1 & cell6 row 1 & cell7 row 1 & cell8 row 1\\ \hline
$cell1 row2$ & cell2 row 2 & cell3 row 2 & cell4 row 2 & cell5 row 2 & cell6 row 2 & cell7 row 2 & cell8 row 2\\ \hline
$cell1 row3$ & cell2 row 3 & cell3 row 3 & cell4 row 3 & cell5 row 3 & cell6 row 3 & cell7 row 3 & cell8 row 3\\ \hline
\end{tabular}
\begin{flushleft} Table notes Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl. Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
\end{flushleft}
\label{table1}
\end{adjustwidth}
\end{table}

%PLOS does not support heading levels beyond the 3rd (no 4th level headings).
\subsection*{\lorem\ and \ipsum\ nunc blandit a tortor}
\subsubsection*{3rd level heading} 
Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque. Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit. Sed non aliquet felis. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Mauris commodo justo ac dui pretium imperdiet. Sed suscipit iaculis mi at feugiat.

% Place figure captions after the first paragraph in which they are cited.
\begin{figure*}[!h]
\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\includegraphics{Fig2}
\caption{{\bf Bold the figure title.}
Figure caption text here, please use this space for the figure panel descriptions instead of using subfigure commands. A: Lorem ipsum dolor sit amet. B: Consectetur adipiscing elit.}
\label{fig2}
\end{adjustwidth}
\end{figure*}

\begin{enumerate}
	\item{react}
	\item{diffuse free particles}
	\item{increment time by dt and go to 1}
\end{enumerate}

\subsection*{Sed ac quam id nisi malesuada congue}

Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam. Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi at justo vitae nulla elementum commodo eu id massa. In vitae diam ac augue semper tincidunt eu ut eros. Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis. Aliquam in enim semper, aliquam massa id, cursus neque. Praesent faucibus semper libero.

\begin{itemize}
	\item First bulleted item.
	\item Second bulleted item.
	\item Third bulleted item.
\end{itemize}

\section*{Discussion}

Top-aligned is not ideal for containment searches.

\section*{Conclusion}

\section*{Supporting information}

% Include only the SI item label in the paragraph heading. Use the \nameref{label} command to cite SI items in the text.
\paragraph*{S1 Fig.}
\label{S1_Fig}
{\bf Bold the title sentence.} Add descriptive text after the title of the item (optional).

\paragraph*{S2 Fig.}
\label{S2_Fig}
{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 File.}
\label{S1_File}
{\bf Lorem ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 Video.}
\label{S1_Video}
{\bf Lorem ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 Appendix.}
\label{S1_Appendix}
{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 Table.}
\label{S1_Table}
{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\section*{Acknowledgments}
Cras egestas velit mauris, eu mollis turpis pellentesque sit amet. Interdum et malesuada fames ac ante ipsum primis in faucibus. Nam id pretium nisi. Sed ac quam id nisi malesuada congue. Sed interdum aliquet augue, at pellentesque quam rhoncus vitae.

\nolinenumbers

% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}
%
% or
%
% Compile your BiBTeX database using our plos2015.bst
% style file and paste the contents of your .bbl file
% here. See http://journals.plos.org/plosone/s/latex for 
% step-by-step instructions.
% 
%\begin{thebibliography}{10}

%\end{thebibliography}

\bibliography{rupee}

\end{document}